Gringo 2.0\\--Draft--

### 2008-09-04 doc/documentation/paper.tex
Doku: removed all /hbox's
git-svn-id: https://potassco.svn.sourceforge.net/svnroot/potassco/branches/gringo2@493 3e02f596-3131-0410-8352-e04eeaec37cc

\paragraph{Predicates and Function Symbols}
Predicates and function symbols have a predicate symbol/name and an arity.
For the function symbol $f$ with arity 3 we write $f/3$.
The names of the function symbols and predicates have to start with either a lower case letter or an underscore and can contain digits and letters.
With $\mathcal{P}$ we denote to the set of predicates and $\mathcal{F}$ denotes the set of function symbols.
\paragraph{Constants}
Constants are function symbols with an arity of zero.
The name of a constant can also be encompassed by double quotation marks, which is sometimes helpful when using constants with spaces. It has to be remarked that a constant with quotation marks is never equal to a constant without, so ``\texttt{a}'' is not equal to \texttt{a}.
$\mathcal{C}$ is the set of constants.
\paragraph{Variables}
Variables have to start with a capitalized letter and can contain digits and letters.
The set of variables is denoted by $\mathcal{V}$.
The language over $\mathcal{P}$, $\mathcal{F}$ and $\mathcal{C}$ is denoted by $\mathcal{L(P,F,C)}$.
\paragraph{Terms}
The building blocks of the language are terms.
Each variable $v \in \mathcal{V}$, constant $c \in \mathcal{C}$ and each function symbol $f \in \mathcal{F}$ is a term.
Furthermore $f(t_1, \dots, t_n)$ is a term if $f/n \in \mathcal{F}$ and $t_i$ for $i \in \{1 \dots n\}$ are terms.
Since a term is defined recursively, it can be nested.
A term with the first parameter being a constant $a$, the others being terms $f(b,X)$ and $g(f(X), Y)$ would be expressed this way.
\begin{verbatim}
 p(a,f(b,X), g(f(X),Y)).
\end{verbatim}
\paragraph{Atoms}
An atom over $\mathcal{L(P, F, C)}$ is of the form $p(t_1, \dots, t_n)$ where $p/n$ is a predicate
symbol with arity $n$, and $t_i \mid i \in \{1 \dots n\}$ are terms.
\paragraph{Rules}
A rule over the language $\mathcal{L(P,F,C)}$ has the form
$$
A_0 \mid \dots \mid A_l \leftarrow A_{l+1}, \dots, A_m, not\ A_{m+1}, \dots, not\ A_n.
$$
where $A_0$ to $A_n$ are atoms. \newline
A logic program $\Pi$ over $\mathcal{L(P,F,C)}$ is a finite set of rules over $\mathcal{L(P,F,C)}$.

@@

 1{move(X,Y,k) : block(X) : location(Y) : X != Y}1.
\end{verbatim}
A move is of the form ``\texttt{move(X,Y,k)}'' where \texttt{X} is the block that is moved, \texttt{Y} the new position of the block and \texttt{k} the step when the moved is executed.
\newline
Now that we have exactly one move per step we need to define what happens if we move a block.
Understandably if we move block \texttt{X} to position \texttt{Y} in step \texttt{k}, the new location of the block shall be \texttt{Y}.
\begin{verbatim}
 on(X,Y,k) :- move(X,Y,k).
\end{verbatim}
For all other blocks that have not moved, the position shall be the same.
In ASP a little trick is used to describe such a condition.
We first define a little helper predicate that states that a block has moved.
\begin{verbatim}
 moved(X, k) :- move(X, Y, k).
\end{verbatim}
With this helping predicate we can define that a block is on the same position in step \texttt{k}, if it was not moved.
\begin{verbatim}
 on(X,Y,k) :- on(X,Y,k-1), block(X), location(Y), not moved(X, k).
\end{verbatim}

@@

\paragraph{\texttt{--ground,-g}}
If this parameter is used, the grounder expects an already grounded program as input.
For converting an already grounded logic program into an output format supported by gringo2.0 this is the right option. This mode saves a lot of memory but does not ground the program.
\paragraph{\texttt{--const,-c C=v}}
You can set the value of constants defined in your logic program via the command line of gringo2.0.
After the parameter ``--const'' you have to give the name of the constant $C$ you want to set followed by a ``$=$'' and the value of that constant. The integer value $v$ can even be an arithmetic expression. For example:
\begin{verbatim}
 gringo --const n=5*2
\end{verbatim}
the constant $n$ is replaced by $10$ in the logic program.
\paragraph{\texttt{--bindersplitting=<arg>}}
With this option you can enable or disable the binder splitting used by gringo2.0
\begin{itemize}
 \item yes : Enable binder splitting
 \item no  : Disable binder splitting
\end{itemize}
\paragraph{\texttt{--ifixed=<num>}}
An incremental program is required as input. Gringo2.0 grounds the fixed number of steps of the incremental program.

\paragraph{Default Parameters}
The default command line is ``\texttt{./gringo --lparse --bindersplitting=yes}'' to read from standard input and for smodels output.

@@

\begin{itemize}
 \item 
\begin{verbatim}
 "Warning: sum with negative bounds in the body of a rule"
\end{verbatim}
This warning occurs if a sum aggregate with negative weights has been found and the output format smodels is used or the output is redirected to clasp. In this format negative bounds are compiled away using the transformation described in \cite{lparseManual}. This may lead to counterintuitive results.
 \item 
\begin{verbatim}
 "Warning: multiple definitions of #const n"
\end{verbatim}
This warning is printed if more than one \#const statement is used to set the same constant.
The last attached value is used for grounding.
\item
\begin{verbatim}
 "Warning: p/3 is never defined"
\end{verbatim}
This warnings states that an atom is used in the body of a rule that never occurs in the head.
The predicate literal is given in the warning. The predicate will be removed through optimization.
\item
\begin{verbatim}
 "got ground program i hope u have enough memory :)"
\end{verbatim}
Use the command line option ``--ground'' to convert a ground program into another format.
For converting already ground programs to an output language a special module has been implemented that uses far less memory. Because trying to ground an already variable free program could need a huge amount of memory.
\item
\begin{verbatim}
 "Warning: There are no #base, #lambda or #delta sections."
\end{verbatim}
If writing an incremental program, the incremental sections\ref{metasyntaxinc} have to be used.
\item
\begin{verbatim}
 "Warning: There are statements not within a #base, #lambda or #delta section."
 "         These Statements are put into the #base section."
\end{verbatim}
If writing an incremental program, all statements that are not written in an incremental section are automatically considered to belong to the \texttt{base} part of the program.
\item
\begin{verbatim}
 "Warning: Classical negation is not handled correctly in combination with the incremental output."
 "         You have to add rules like: :- a, -a. on your own! (at least for now)"
\end{verbatim}
To use classical negation in an incremental program the described rules have to be added manually. This will be fixed in a later version of gringo.
\end{itemize}
Furthermore all warnings from clasp are included.

\subsection{Errors}
gringo2.0 stops the execution and prints an error message to standard error output.
\begin{itemize}
 \item 
\begin{verbatim}
 "the following rule cannot not be grounded,
  non domain predicates : { ... }"
\end{verbatim}
This error message gives a list of predicates that are not domain restricted.
You explicitly have to give a domain for those predicates.
 \item 
\begin{verbatim}
 "the following rule cannot not be grounded,
  weakly restricted variables: { ... }"
\end{verbatim}
This error message gives a list of variables that are not lambda restricted. 
\item
\begin{verbatim}
 "syntax error on line Y column Z unexpected token: X"
\end{verbatim}
The parser can not read the input format. Some unexpected token occurred in the input file.
\item
\begin{verbatim}
 "sorry clasp cannot handle disjunctive rules!"
\end{verbatim}
If using clingo or iclingo, this happens if disjunctive rules are used. Please choose another solver that support disjunctive rules.
\item
\begin{verbatim}
 "A fixed number of incremental steps is needed to ground the program."
\end{verbatim}
This error occurs if an incremental program is given to gringo without using the option \texttt{--ifixed} or iclingo.
\item
\begin{verbatim}
 "cyclic constant definition."
\end{verbatim}
A cyclic definition of a constant is not allowed as shown in the example.
\begin{verbatim}
 #const x=y.
 #const y=x.
\end{verbatim}
\item
\begin{verbatim}
 "The following statement cant be used with the incremental interface: X."
 "unsupported statement."
\end{verbatim}
It is not allowed to use an optimize (\texttt{minimize} or \texttt{maximize}) or \texttt{compute} statement in an incremental program.
\item
\begin{verbatim}
 "comparing different types"
 "comparing function symbols"
\end{verbatim}
This error occurs if trying to compare non compareable values or function symbols as in
\begin{verbatim}
 p :- 1 < s.
 p :- f(x) < f(y).
\end{verbatim}

>>>>

\paragraph{Predicates and Function Symbols}
Predicates and function symbols have a predicate symbol/name and an arity.
For the function symbol $f$ with arity 3 we write $f/3$.
The names of the function symbols and predicates have to start with either a lower case letter or an underscore and can contain digits and letters.
With $\mathcal{P}$ we denote to the set of predicates and $\mathcal{F}$ denotes the set of function symbols.
\paragraph{Constants}
Constants are function symbols with an arity of zero.
The name of a constant can also be encompassed by double quotation marks, which is sometimes helpful when using constants with spaces. It has to be remarked that a constant with quotation marks is never equal to a constant without, so ``\texttt{a}'' is not equal to \texttt{a}.
$\mathcal{C}$ is the set of constants.
\paragraph{Variables}
Variables have to start with a capitalized letter and can contain digits and letters.
The set of variables is denoted by $\mathcal{V}$.
The language over $\mathcal{P}$, $\mathcal{F}$ and $\mathcal{C}$ is denoted by $\mathcal{L(P,F,C)}$.
\paragraph{Terms}
The building blocks of the language are terms.
Each variable $v \in \mathcal{V}$, constant $c \in \mathcal{C}$ and each function symbol $f \in \mathcal{F}$ is a term.
Furthermore $f(t_1, \dots, t_n)$ is a term if $f/n \in \mathcal{F}$ and $t_i$ for $i \in \{1 \dots n\}$ are terms.
Since a term is defined recursively, it can be nested.
A term with the first parameter being a constant $a$, the others being terms $f(b,X)$ and $g(f(X), Y)$ would be expressed this way.
\begin{verbatim}
 p(a,f(b,X), g(f(X),Y)).
\end{verbatim}
\paragraph{Atoms}
An atom over $\mathcal{L(P, F, C)}$ is of the form $p(t_1, \dots, t_n)$ where $p/n$ is a predicate
symbol with arity $n$, and $t_i$ for $i \in \{1 \dots n\}$ are terms.
\paragraph{Rules}
A rule over the language $\mathcal{L(P,F,C)}$ has the form
$$
A_0 \mid \dots \mid A_l \leftarrow A_{l+1}, \dots, A_m, not\ A_{m+1}, \dots, not\ A_n.
$$
where $A_0$ to $A_n$ are atoms. \newline
A logic program $\Pi$ over $\mathcal{L(P,F,C)}$ is a finite set of rules over $\mathcal{L(P,F,C)}$.

@@

 1{move(X,Y,k) : block(X) : location(Y) : X != Y}1.
\end{verbatim}
A move is of the form ``\texttt{move(X,Y,k)}'' where \texttt{X} is the block that is moved, \texttt{Y} the new position of the block and \texttt{k} the step when the moved is executed.
\newline
Now that we have exactly one move per step we need to define what happens if we move a block.
Understandably if we move block \texttt{X} to position \texttt{Y} in step \texttt{k}, the new location of the block shall be \texttt{Y}.
\begin{verbatim}
 on(X,Y,k) :- move(X,Y,k).
\end{verbatim}
For all other blocks that have not moved, the position shall be the same.
In ASP a little trick is used to describe such a condition.
We first define a little helper predicate that states that a block has moved.
\begin{verbatim}
 moved(X, k) :- move(X, Y, k).
\end{verbatim}
With this helping predicate we can define that a block is on the same position in step \texttt{k}, if it was not moved.
\begin{verbatim}
 on(X,Y,k) :- on(X,Y,k-1), block(X), location(Y), not moved(X,k).
\end{verbatim}

@@

\paragraph{\texttt{--ground,-g}}
If this parameter is used, gringo expects an already grounded program as input.
For converting an already grounded logic program into an output format 
supported by gringo2.0 this is the right option.
 This mode saves a lot of memory but does not ground the program.
\paragraph{\texttt{--const,-c C=v}}
You can set the value of constants defined in your logic program via the command line of gringo2.0.
After the parameter ``--const'' you have to give the name of the constant $C$ you want to set followed by a ``$=$'' and the value of that constant. The integer value $v$ can even be an arithmetic expression. For example:
\begin{verbatim}
 gringo --const n=5*2
\end{verbatim}
the constant $n$ is replaced by $10$ in the logic program.
\paragraph{\texttt{--bindersplitting=<arg>}}
With this option you can enable or disable the binder splitting used by gringo2.0
\begin{itemize}
 \item yes : Enable binder splitting
 \item no  : Disable binder splitting
\end{itemize}
\paragraph{\texttt{--ifixed=<num>}}
An incremental program is required as input. Gringo2.0 grounds the fixed number of steps of the incremental program.

\paragraph{Default Parameters}
The default command line is
\begin{verbatim}
 gringo --lparse --bindersplitting=yes
\end{verbatim}
to read from standard input and for smodels output.

@@

\begin{itemize}
 \item 
\begin{verbatim}
 "Warning: sum with negative bounds in the body of a rule"
\end{verbatim}
This warning occurs if a sum aggregate with negative weights has been found and the output format smodels is used or the output is redirected to clasp. In this format negative bounds are compiled away using the transformation described in \cite{lparseManual}. This may lead to counterintuitive results.
 \item 
\begin{verbatim}
 "Warning: multiple definitions of #const n"
\end{verbatim}
This warning is printed if more than one \#const statement is used to set the same constant.
The last attached value is used for grounding.
\item
\begin{verbatim}
 "Warning: p/3 is never defined"
\end{verbatim}
This warnings states that an atom is used in the body of a rule that never occurs in the head.
The predicate literal is given in the warning. The predicate will be removed through optimization.
\item
\begin{verbatim}
 "got ground program i hope u have enough memory :)"
\end{verbatim}
Use the command line option ``--ground'' to convert a ground program into another format.
For converting already ground programs to an output language a special module has been implemented that uses far less memory. Because trying to ground an already variable free program could need a huge amount of memory.
\item
\begin{verbatim}
 "Warning: There are no #base, #lambda or #delta sections."
\end{verbatim}
If writing an incremental program, the incremental sections\ref{metasyntaxinc} have to be used.
\item
\begin{verbatim}
 "Warning: There are statements not within a #base, #lambda"
 "         or #delta section."
 "         These Statements are put into the #base section."
\end{verbatim}
If writing an incremental program, all statements that are not written in an incremental section are automatically considered to belong to the \texttt{base} part of the program.
\item
\begin{verbatim}
 "Warning: Classical negation is not handled correctly in"
 "         combination with the incremental output."
 "         You have to add rules like:"
 "         :- a, -a. on your own! (at least for now)"
\end{verbatim}
To use classical negation in an incremental program the described rules have to be added manually. This will be fixed in a later version of gringo.
\end{itemize}
Furthermore all warnings from clasp are included.

\subsection{Errors}
gringo2.0 stops the execution and prints an error message to standard error output.
\begin{itemize}
 \item 
\begin{verbatim}
 "the following rule cannot not be grounded,
  non domain predicates : { ... }"
\end{verbatim}
This error message gives a list of predicates that are not domain restricted.
You explicitly have to give a domain for those predicates.
 \item 
\begin{verbatim}
 "the following rule cannot not be grounded,
  weakly restricted variables: { ... }"
\end{verbatim}
This error message gives a list of variables that are not lambda restricted. 
\item
\begin{verbatim}
 "syntax error on line Y column Z unexpected token: X"
\end{verbatim}
The parser can not read the input format. Some unexpected token occurred in the input file.
\item
\begin{verbatim}
 "sorry clasp cannot handle disjunctive rules!"
\end{verbatim}
If using clingo or iclingo, this happens if disjunctive rules are used. Please choose another solver that support disjunctive rules.
\item
\begin{verbatim}
 "A fixed number of incremental steps is needed"
 "to ground the program."
\end{verbatim}
This error occurs if an incremental program is given to gringo without using the option \texttt{--ifixed} or iclingo.
\item
\begin{verbatim}
 "cyclic constant definition."
\end{verbatim}
A cyclic definition of a constant is not allowed as shown in the example.
\begin{verbatim}
 #const x=y.
 #const y=x.
\end{verbatim}
\item
\begin{verbatim}
 "The following statement cant be used with the"
 "incremental interface: X."
 "unsupported statement."
\end{verbatim}
It is not allowed to use an optimize (\texttt{minimize} or \texttt{maximize}) or \texttt{compute} statement in an incremental program.
\item
\begin{verbatim}
 "comparing different types"
 "comparing function symbols"
\end{verbatim}
This error occurs if trying to compare non compareable values or function symbols as in
\begin{verbatim}
 p :- 1 < s.
 p :- f(x) < f(y).
\end{verbatim}
